---
title: "Regression Assignment"
author: "Venkata Naga Siddartha Gutha"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. a)
```{r}
set.seed(2017) 
X=runif(100)*10 
Y=X*4+3.45 
Y=rnorm(100)*0.29*Y+Y 
plot(Y~X)

```

from the graph it can bee seen that value of y increases with incease in x.This indicates that there is some relation between X and Y. So, yes they can be put into a linear model to explain y based on x


1. b)
```{r}
model<-lm(Y~X)
model
summary(model)


```
Equation:
Y= 4.465+(3.611)*X


Accuracy of a model is determained by r Squared value and in this case it is 0.6517, this means that the model explains 65.17% variability of the response variable i.e Y.

1. c)


In the above case Correlation Coefficient is equal to the Coefficent of Determination (R^2 ) as the regression is based on single variable. Therefore Correlation Coefficient = R^2=O.6517


2. a)
```{r}
head(mtcars)
lm_james<-lm(hp~wt,data = mtcars)
summary(lm_james)

lm_chris<-lm(hp~mpg,data = mtcars)
summary(lm_chris)


```
Chris is correct as the model made on Chris opinion has an accuracy of 0.6024 which is higher than accuracy of 0.4339 of the model made on James opinion 

2. b)
```{r}
model_hp<-lm(hp~cyl+mpg,data = mtcars)
summary(model_hp)

predict(model_hp,data.frame(cyl=4,mpg=22))


```

From the model it is estimated that the Horse Power of a car with 4 calendar and mpg of 22 is 88.93618.


3. a)
```{r}
library(mlbench)
data("BostonHousing")
head(BostonHousing)

Boston_model<-lm(medv~crim+zn+ptratio+chas,data = BostonHousing)
summary(Boston_model)
```

From the summary of the model it can be seen r squared value is 0.3588. This implies that accuracy of the model is low.



3. b) i.

When two houses are identical in all aspects then the one that bounds Charles River is more expensive by $4583.9. This is because coefficient of Charles variable is 4.5839 and price is measured inn 1000's of USD.



3. b) ii.

Coefficient of pupil-teacher ratio from the model is -1.493. This indicates that for every unit increase of pupil-teacher ratio the price of house decreases by $1493. So, the price of house with pupil-teacher ratio 15, 18 is reduced by 22395 and 26874 USD respectively. Therefore the house with pupil_teacher ratio 15 is USD 4479 more expensive than the house with pupil_teacher ratio 18.



3. c)


The variables crime  crate  (crim),  proportion  of  residential  land zoned  for  lots  over  25,000  sq.ft  (zn),  the  local pupil-teacher ratio  (ptratio)  and  whether  the  tract  bounds Charles River(chas) that are used in the model are all statistically important as their p-values lies between 0 and 0.001.



3. d)
```{r}
anova(Boston_model)

```
From the anova analysis the order of importance of the four variables is as follows:


1.crime  crate  (crim)


2.the  local pupil-teacher ratio  (ptratio)


3.proportion  of  residential  land zoned  for  lots  over  25,000  sq.ft  (zn)


4.the  tract  bounds Charles River(chas)

```{r}


```


```{r}


```


```{r}


```

